Efficient Parallelization of $\ell_1$-Minimization for Face Recognition \thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}

### 2011-02-14 paper/main.tex
Diagram getting included properly.
Alignment and font still need work.
One solution might be to typeset
a full-size version of each diagram
with appropriately scaled font, convert
to ps, and include that.
Re-wrote part of introduction, Simplified Diagram

%Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization.
$\ell_1$-minimization is one of the most powerful tools to be added to the engineer's
toolbox in years.  It has strong theoretical motivation for both for use in recovering sparse
solutions to systems of linear equations, and has been shown to perform well even in the
presence of a combination of sparse and dense error.  Since solutions of linear systems
arise in practically every branch of engineering and there are very few application which
do {\em not} benefit from a robust error function, the applications are far to numerous to
itemize here.  
For many applications of $\ell_1$-minimization solving a system $b=Ax$, $A$ often has a
special structure that can be leveraged to compute portions of $A$ on the fly
as they are needed, speeding up perfomance of the minimization dramatically.
Unfortunately, there are many import applications where this does not hold true.
This paper targets applications where $\bb$ and $A$ contain
image data, and thus must be loaded every time they are used, as is the 
case for many problems in computer vision.
The techniques presented in this paper target these applications, as well
as others where the dictionary is dense and unstructured. $A$ must be loaded every time
it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
than arithmetic logic.

% Narrow focus for applications to Recognition Application
The primary motivating application for this paper is face recognition for access control applications.
Recently, advances in automatic face recognition have been made by re-casting
it as a sparse representation problem.  This core of this technique consists of
re-casting recognition as a sparse representation problem. This is done by  
stacking the test image into a vector $\bb$, stacking the training images for
all of the users into the columns of a matrix $A$, and solving the following
minimization problem:
\begin{equation}
\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.

For this idea to work, all of the images must be aligned. An iterative alignment
algorithm for this purpose was described as part of the \cite{Wagner2009-CVPR}, which was
the first paper to expand \cite{Wright2009-PAMI} into a functioning face recognition system.
This iterative alignment routine is based on repeatedly solving the following optimization problem:
\begin{equation}
\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_k \x + \e.
\end{equation}
Where $A_k$ contains the training images for user $k$ as well as several vectors consisting of the
Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
both the current representation coefficients, as well as an update to the transformation parameters.

\begin{figure}
\centering
%\includegraphics[width=3.5in]{figures/pipeline.pdf}
\input{figures/pipeline.tex}
\caption{The Full Recognition Pipeline}
\label{fig:pipeline}
\end{figure}

>>>>

%Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization.
$\ell_1$-minimization is one of the most powerful tools to be added to the engineer's
toolbox in years.  It has strong theoretical motivation for both for use in recovering sparse
solutions to systems of linear equations, and has been shown to perform well even in the
presence of a combination of sparse and dense error.  Since solutions of linear systems
arise in practically every branch of engineering and there are very few application which
do {\em not} benefit from a robust error function, the applications are far to numerous to
itemize here.  
For many applications of $\ell_1$-minimization solving a system $b=Ax$, $A$ often has a
special structure that can be leveraged to compute portions of $A$ on the fly
as they are needed, speeding up perfomance of the minimization dramatically.
Unfortunately, there are many import applications where this does not hold true.
This paper targets applications where $\bb$ and $A$ contain
image data, and thus must be loaded every time they are used, as is the 
case for many problems in computer vision.
The techniques presented in this paper target these applications, 
but are relevant to other non-vision applications where the dictionary is dense and unstructured. 
%$A$ must be loaded every time
%it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
%than arithmetic logic.

The primary motivating application for this paper is face recognition for access control applications.
Recently, advances in automatic face recognition have been made by re-casting
it as a sparse representation problem.  This core of this technique consists of
re-casting recognition as a sparse representation problem. This is done by  
stacking the test image into a vector $\bb$, stacking the training images for
all of the users into the columns of a matrix $A$, and solving the following
minimization problem:
\begin{equation}
\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.

For this idea to work, all of the images must be aligned. An iterative
alignment algorithm for this purpose was described described in
\cite{Wagner2009-CVPR}.  The algorithm essentially extends Lucas-Kanade
iterative alignment to use the $\ell_1$ norm as a robust error function, and to
simultaneously estimate the illumination in terms of the gallery illuminations.
In combination with a custom training image acquisition system, this resulted
in a complete face recognition system that achieved good recognition rates even
on data captured in realistic situations.  Unfortunately, the computational
cost of this algorithm was still too high for production use.  This speed was
somewhat improved in \cite{Wagner2010-PAMI}, which showed improved performance
by using Augmented Lagrange Multiplier (ALM) based $\ell_1$ minimizations
rather than interior point methods.  This brought alignment time down to about
half a second per training user, this is still far too slow to be useful in
access control scenariors where the gallery contains hundreds of users.
The primary contribution of this paper is to demonstrate how massively
parallel hardware can dramatically improve recognition time in this application,
finally making it practical for real-world use.

We now provide an overview of the recognition pipeline proposed 
in \cite{Wagner2010-PAMI}.  The full recognition pipeline can be divided into 5 major stages.

(... I will describe the stages in more detail with equations...)

These stages are summarized in the diagram in Figure \ref{fig:pipeline}.

%This iterative alignment routine is based on repeatedly solving the following optimization problem:
%\begin{equation}
%\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_i \x + \e.
%\end{equation}
%Where $A_i$ contains the training images for user $i$ as well as several vectors consisting of the
%Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
%both the current representation coefficients, as well as an update to the transformation parameters.

\begin{figure}
\centering
{\tiny \input{figures/pipeline_simplified.tex}}
\caption{The Full Recognition Pipeline}
\label{fig:pipeline}
\end{figure}

\begin{algorithm}[thb]
\caption{\bf (Deformable Sparse Recovery and Classification for
Face Recognition)} \label{alg:deformable-src}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\bb\in\Re^m$ and a deformation group $T$. 
\STATE
{\bf for} each subject $i$, 
\STATE \hspace{3mm} $\tau^{(0)}
\leftarrow I$. 
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
\STATE \hspace{6mm}
$\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ
\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
\tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \bb + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
\subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
\STATE
\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$; 
\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
residuals $\|\e\|_1$. \STATE Compute an average transformation
$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
\STATE Update $\bb \leftarrow \bb \circ \bar{\tau}$ and $\tau_i
\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
$\ell^1$-minimization problem: \hspace{2mm}
$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \bb = A \x + \e.$
\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\bb) = \arg\min_i r_i(\y)$.
\end{small}
\end{algorithmic}
\end{algorithm}
